---
title: "Exploratory Data Analysis"
author: "Chindu"
date: "5/26/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This is a Expploratory Data Analysis report carried out on a sample CRM dataset


```{r,include= FALSE}
library(dplyr)
library(dlookr)
```

Loading csv file into R studio
```{r}
data<-read.csv("ReadyforModelling.csv") 
```

Checking if R studio has identified the right structure for each variable
```{r}
str(data)
```
Changing structure of wrongly assigned variables and remove variables unrealated to the analysis
```{r}
data$Answered.by.specialist<- factor(data$Answered.by.specialist)
data$Booked.Status<- factor(data$Booked.Status)
data$EnquiryYear<-factor(data$EnquiryYear)
data$DepYear<-factor(data$DepYear)
data$X<-NULL
```
Get a better understanding of the entire dataset
```{r}
summary(data)
```
Get a better understanding of numeric/interger variables
```{r}
diagnose_numeric(data)
```
Get a better understanding of categorical variables
```{r}
diagnose_category(data)
```
Checking correlation between numerical variables
```{r}
plot_correlate(data)
```
Exploring relation between target varaible (Booked.Status) and a numeric variable
```{r}
categ<-target_by(data,Booked.Status)
cat_num<-relate(categ,Duration)
cat_num
summary(cat_num)
plot(cat_num) # relationship between booked.status and duration is represented using a desity plot
```
Exploring relation between target variable(BookedStatus) and a categorical variable
```{r}
cat_cat<-relate(categ,Allocated.Time)
cat_cat
plot(cat_cat) #mosaics plot
```



